---
title: "Repeatable Modelling Workflow"
author: "Kim Cressman"
date: "created 2022-02-08; latest update `r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
```


This document pulls together all the model-running and coefficient-presenting code, by using previous scripts as child documents.  

# Data Import and Summaries  

```{r, child = '03_starting_over.Rmd'}

```


# Exploring Possible Transformations of Response  

Fitting a linear model with all possible predictors, then evaluating residuals and predictive performance in a rough way to see which transformation works best. This exploration led to the decision to use a square-root transformation and all following model selection will be based on square-root transformation of the response variable (extracted chlorophyll).  

```{r, child = '04_new_models_and_transformations.Rmd'}

```


# Generating statistical models  

