---
title: "Repeatable Modelling Workflow"
author: "Kim Cressman"
date: "created 2022-02-08; latest update `r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}
library(readxl)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(lmerTest)
library(MuMIn)
library(boot)
library(tibble)

# cost function for sMdAPE (symmetric median absolute percent error)  
# works on square-root transformed response variable  
cost <- function(obs, expt){
    median(
        abs(
            100 * (obs^2-expt^2) / ((obs^2+expt^2)/2)
        )
    )
}
```


This document pulls together all the model-running and coefficient-presenting code, by using previous scripts as child documents.  

# Data Import and Summaries  

A few notes.  

+  NIW collected tank samples from one site and ISCO samples from a different site. Due to their concerns about pooling data from the two sites (they are very different environments), I created a 'NIW_isco' reserve ID to differentiate.  
+  For a similar reason, I created a 'HEE_w2' reserve ID for the site where He'eia started getting samples for this project, but the water is so clear that the sensor wasn't working well. Figured I'd separate it out.  
+  GRB used 8 different stations throughout. *I did not differentiate.* If models fit poorly for this reserve, this could be a reason.  
+  PDB used 3 differents stations, and again I did not differentiate.  
+  I made up a station ID for Sapelo because they did not use a normal SWMP station. I called it 'sapnawq' (for 'N/A', ha)  

```{r, child = '03_starting_over.Rmd'}

```


# Exploring Possible Transformations of Response  

Fitting a linear model with all possible predictors, then evaluating residuals and predictive performance in a rough way to see which transformation works best. This exploration led to the decision to use a square-root transformation and all following model selection will be based on square-root transformation of the response variable (extracted chlorophyll).  

```{r, child = '04_new_models_and_transformations.Rmd'}

```


Initial data and model exploration (not shown in this document; see instead the older `01_first_models_and_transformations` files) suggested that distinguishing between ISCO and tank samples at a site did not improve the model (no better AIC than lumping). However, due to huge anecdotal differences at two reserves (GTM and LKS) and the potential for differences at other reserves, models will be run both ways: with all data (Tank + ISCO), and with tank-only data.    

# Models + Graphics: All Reserves Combined; Tank + ISCO  


```{r, child = '09b_minimal_combinedReserves_TankandISCO2.Rmd'}


```


# Models + Graphics: All Reserves Combined; Tank Only  

```{r, child = '09b_minimal_combinedReserves_TankOnly.Rmd'}

```


# Models: Individual Reserves; Tank + ISCO  

```{r, child = '10_minimal_indRes_TankAndISCO.Rmd'}

```

# Models: Individual Reserves; Tank Only  

```{r, child = '10_minimal_indRes_TankOnly.Rmd'}

```

# Graphic: R^2 and prediction error for all models  

Reserve code 'ALL' represents the combined-reserve models.  

```{r, child = '11_summary_diagnostic_graphic.Rmd'}

```

