---
date: "12/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Having decided in the previous file (`01_first_models_and_transformations`) that a square-root transformation of the response was most appropriate, all models in this file will be based on such a transformation.  

In that file I did not see a difference between model outputs between the predictors `sensor_ugl` vs. `sensor_rfu`, so to fit with previous conversations by the workgroup, I'll proceed using only `sensor_rfu`.  

Finally, there was very little difference in model fit when including `method` (tank vs. ISCO) as when excluding it, so we do not need to include it as a variable. AIC was a bit improved without it.  

The rest of this file will focus on other variable selection.  

```{r, warning = FALSE, message = FALSE}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
```

```{r}
dat <- read_xlsx(here::here("data", "2021_chla-catalyst_data_all.xlsx"),
                 sheet = "qaqc") %>% 
    mutate(extracted = chla_ugl,
           sensor_rfu = chlorophyll_rfu,
           sensor_ugl = chl_fluor,
           month = lubridate::month(datetime_collected)) %>% 
    select(reserve = reserve_code,
           sample_no,
           station_code,
           datetime_collected,
           month,
           extracted,
           sensor_rfu,
           temp,
           sal,
           turb,
           fdom_qsu,
           do_mgl,
           do_pct) %>% 
    filter(!is.na(extracted),
           !is.na(sensor_rfu))  # can probably work around other NAs, but HAVE to have both chlorophyll measurements  

summary(dat)
nrow(dat)
```

# Full model  

...excluding DO in all preliminary things, might add in later to see? But this is the full model that gave me the best (*most okay*) fit in the previous file.

sqrt(chl_extracted) ~ sensor_rfu + turbidity + fdom + turb*fdom + temp + salinity + month + reserve

```{r}
fit_full <- lm(sqrt(extracted) ~ sensor_rfu + turb + fdom_qsu + turb*fdom_qsu + temp + sal + as.factor(month) + as.factor(reserve), data = dat)

broom::glance(fit_full)


all_big <- dat %>% 
    filter(!is.na((extracted & sensor_rfu & turb & 
                       fdom_qsu & temp & sal)))

all_big$fitted <- fitted.values(fit_full)

ggplot(all_big) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(aes(x = fitted, y = extracted, col = reserve)) +
    facet_wrap(~month, ncol = 4) +
    labs(x = "fitted",
         y = "observed",
         title = "All months",
         subtitle = paste(fit_full$call[2],
                          "\nAdjusted R^2:",
                          round(summary(fit_full)$adj.r.squared, 3),
                          "\nAIC:",
                          round(AIC(fit_full), 2)))

```

From the different viewing I did of models in the `01` file, it looked like the turb*FDOM interaction could be dropped from the model. So I'll do a few models without that.  

I'd also love to drop reserve and month. Let's see if dropping those three things still gets us a reasonable fit.  

# Reduced 1 - drop only the turb*fdom interaction  

```{r}
fit_red1 <- lm(sqrt(extracted) ~ sensor_rfu + turb + fdom_qsu + temp + sal + as.factor(month) + as.factor(reserve), data = dat)

broom::glance(fit_red1)
```

That's about as good and even better by a couple of AIC points. So I won't bring the interaction in and out; we'll just leave it out moving forward.  

# Reduced 2 - drop temp and sal; keep month and reserve  


```{r}
fit_red2 <- lm(sqrt(extracted) ~ sensor_rfu + turb + fdom_qsu + as.factor(month) + as.factor(reserve), data = dat)

broom::glance(fit_red2)
```

A little worse, but not much.    

# Reduced 3 - drop month too  

```{r}
fit_red3 <- lm(sqrt(extracted) ~ sensor_rfu + turb + fdom_qsu + as.factor(reserve), data = dat)

broom::glance(fit_red3)
```
  
month turns out to be important


# Reduced 4 - month is back; drop reserve  

```{r}
fit_red4 <- lm(sqrt(extracted) ~ sensor_rfu + turb + fdom_qsu + as.factor(month), data = dat)

broom::glance(fit_red4)
```

Noooope. Reserve needs to stay in.  


# Reduced 5 - drop FDOM  

Please, we can save $4k if we don't need FDOM! Month and Reserve are in this one because they seem to matter.  

```{r}
fit_red5 <- lm(sqrt(extracted) ~ sensor_rfu + turb +  as.factor(month) + as.factor(reserve), data = dat)

broom::glance(fit_red5)
```

Adjusted R^2 isn't bad, but AIC got worse. This might not be a clear answer.  

......and it may depend on reserve:  

# Reduced-ish 6: include FDOM*Reserve interaction  

```{r}
fit_red6 <- lm(sqrt(extracted) ~ sensor_rfu + turb + fdom_qsu + as.factor(reserve) + fdom_qsu*as.factor(reserve), data = dat)

broom::glance(fit_red6)
```

Not the greatest. Not the worst. Probably have to evaluate on a reserve-level basis.  

# Reduced 7 - only reserve/month, not turb/fdom?  

I'd be surprised if this isn't terrible.....

```{r}
fit_red7 <- lm(sqrt(extracted) ~ sensor_rfu + as.factor(month) + as.factor(reserve), data = dat)

broom::glance(fit_red7)
```

Hm. That adjusted R^2 is almost as good as the original. AIC is a fair amount worse.  

I think this is all pointing to FDOM - and possibly turbidity - as reserve-dependent predictors.  