---
output: 
    html_document
---

**In this file**   

+  Data will be subsetted to a data frame that does not contain ANY NAs, because model selection is based on models containing exactly the same input data.  
+  That subset will be saved as an `.Rdata` file for easier import and reuse downstream.  
+  Some summary statistics for the data that will be used will be generated.  
## Data import and export  

re-run 12/16/2021 to incorporate new data file sent on 12/15 (GRB ISCO data had been left out of original file)  

```{r}
dat <- read_xlsx(here::here("data", "2021_chla-catalyst_data_all.xlsx"),
                 sheet = "qaqc") %>% 
    mutate(extracted = chla_ugl,
           sensor_rfu = chlorophyll_rfu,
           sensor_ugl = chl_fluor,
           month = lubridate::month(datetime_collected),
           season = case_when(month %in% c(12, 1, 2)  ~ "Winter",
                              month %in% c(3, 4, 5)   ~ "Spring",
                              month %in% c(6, 7, 8)   ~ "Summer",
                              month %in% c(9, 10, 11) ~ "Fall")) %>% 
    select(reserve = reserve_code,
           station = station_code,
           datetime_collected,
           month,
           season,
           method,
           extracted,
           sensor_rfu,
           temp,
           turb,
           fdom_qsu) %>% 
    mutate(station = case_when(reserve == "GND" ~ "gndblwq",
                               reserve == "SAP" ~ "sapnawq",
                               TRUE ~ station),
           reserve = case_when(station == "heew2wq" ~ "HEE_w2",
                               station == "WS" ~ "NIW_isco", # at NIW: OL is tank; WS is ISCO. These should not be pooled.
                               TRUE ~ reserve),
           station = case_when(station == "OL" ~ "niwolwq",
                               station == "WS" ~ "niwwswq",
                               TRUE ~ station)) %>% 
    drop_na() %>% 
    mutate(across(c(reserve, month, season),
                  as.factor),
           season = forcats::fct_relevel(season, c("Winter", "Spring", "Summer", "Fall")))

# save(dat, file = here::here("data", "subset_no_NA.RData"))

# read back in later with:
# load(here::here("data", "subset_no_NA.RData"))
```

## Summary  

### General  

```{r}
dat %>% group_by(reserve) %>% count(station) %>% knitr::kable()
summary(dat)
nrow(dat)
```


### Reserve by Month  

Seems like some of us didn't get the temporal spread we were wanting.  


```{r}
table(dat$reserve, dat$month)
```

```{r}
dat %>% 
    group_by(reserve, month, method) %>% 
    summarize(n = n()) %>% 
    ggplot(aes(x = method, y = n, col = reserve)) +
    geom_point(size = 2, alpha = 0.6, position = position_dodge2(width = 0.3)) +
    facet_wrap(~month, ncol = 4) +
    labs(title = "# Samples by Month, by Reserve") +
    theme_bw()
```

### Reserve by Season  

```{r}
table(dat$reserve, dat$season)
```

```{r}
dat %>% 
    group_by(reserve, season, method) %>% 
    summarize(n = n()) %>% 
    ggplot(aes(x = method, y = n, col = reserve)) +
    geom_point(size = 2, alpha = 0.6, position = position_dodge2(width = 0.3)) +
    facet_wrap(~season, ncol = 2) +
    labs(title = "# Samples by Season, by Reserve") +
    theme_bw()
```


Values by season and reserve.....

```{r}
dat %>% 
    group_by(reserve, season, method) %>% 
    ggplot(aes(x = method, y = extracted, col = reserve)) +
    geom_point(size = 2, alpha = 0.6, position = position_dodge2(width = 0.3)) +
    facet_wrap(~season, ncol = 2) +
    labs(title = "Extracted chl a by Season, by Reserve") +
    theme_bw()
```


```{r}
dat %>% 
    group_by(reserve, season, method) %>% 
    ggplot(aes(x = method, y = fdom_qsu, col = reserve)) +
    geom_point(size = 2, alpha = 0.6, position = position_dodge2(width = 0.3)) +
    facet_wrap(~season, ncol = 2) +
    labs(title = "FDOM (qsu) by Season, by Reserve") +
    theme_bw()
```


### Correlations between predictors?  

```{r, fig.width = 8, fig.height = 8}
dat %>% 
    select(reserve, month, season, method, extracted, sensor_rfu, temp, turb, fdom_qsu) %>% 
    plot()
```

