---
date: "12/8/2021"
output: 
    html_document:
        code_folding: hide
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

The previous files had some extensive EDA. I did not, however, include interactions between `sensor_rfu` and anything else, adn this may be important. May also change some things, but might not.  

+  Square-root transformation of the response variable (`extracted`) seemed most appropriate (see `01_first_models_and_transformations`). This will be reexamined.      
+  In that file I did not see a difference between model outputs between the predictors `sensor_ugl` vs. `sensor_rfu`, so to fit with previous conversations by the workgroup, I'll proceed using only `sensor_rfu`. This should still be valid even though interactions weren't included.  
+  Finally, there was very little difference in model fit when including `method` (tank vs. ISCO) as when excluding it, so we do not need to include it as a variable. AIC was a bit improved without it. Again, this did not include interactions, but should still be valid.  

**In this file**   

+  Data will be subsetted to a data frame that does not contain ANY NAs, because model selection is based on models containing exactly the same input data.  
+  That subset will be saved as an `.Rdata` file for easier import and reuse downstream.  
+  Some summary statistics for the data that will be used will be generated.  


```{r, warning = FALSE, message = FALSE}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
```

# Data import and export  

```{r}
dat <- read_xlsx(here::here("data", "2021_chla-catalyst_data_all.xlsx"),
                 sheet = "qaqc") %>% 
    mutate(extracted = chla_ugl,
           sensor_rfu = chlorophyll_rfu,
           sensor_ugl = chl_fluor,
           month = lubridate::month(datetime_collected)) %>% 
    select(reserve = reserve_code,
           sample_no,
           station_code,
           datetime_collected,
           month,
           method,
           extracted,
           sensor_rfu,
           temp,
           turb,
           fdom_qsu) %>% 
    drop_na()   

# save(dat, file = here::here("data", "subset_no_NA.RData"))

# read back in later with:
# load(here::here("data", "subset_no_NA.RData"))
```

# Summary  

## General  

```{r}
summary(dat)
nrow(dat)
```


## Reserve by Month  

Seems like some of us didn't get the temporal spread we were wanting.  


```{r}
table(dat$reserve, dat$month)
```

```{r}
dat %>% 
    group_by(reserve, month, method) %>% 
    summarize(n = n()) %>% 
    ggplot(aes(x = method, y = n, col = reserve)) +
    geom_point(size = 2, alpha = 0.6, position = position_dodge2(width = 0.3)) +
    facet_wrap(~month, ncol = 4) +
    labs(title = "# Samples by Month, by Reserve") +
    theme_bw()
```

