---
output: 
    html_document
---

**In this file**   

+  Data will be subsetted to a data frame that does not contain ANY NAs, because model selection is based on models containing exactly the same input data.  
+  That subset will be saved as an `.Rdata` file for easier import and reuse downstream.  
+  Some summary statistics for the data that will be used will be generated.  
## Data import and export  

re-run 12/16/2021 to incorporate new data file sent on 12/15 (GRB ISCO data had been left out of original file)  

**update 2/16/2022 to incorporate temperature-corrected RFU** (based on Hannah Ramage's calculations) into the modeling data frame. The "global" rho is an average of the rho values in all of the temperature-interference trials. The global rho is then inserted into the Watras et al. 2017 Eq. 1:  

Temp-corrected-RFU = original-RFU / (1 + global_rho*(temp_reading - 20))  


I'm hardcoding in the rhos here because it's easier than dealing with another file, but am setting up the code to be rerun easily with a different batch of rhos (however they're read in).  

```{r}
rhos <- c(-0.018449239,
-0.017224308,
-0.006810888,
-0.017013983,
-0.016389726,
-0.016809971,
-0.010759556,
-0.011669893,
-0.006993083,
-0.010889714,
-0.0081596,
-0.002334279
)

global_rho <- mean(rhos)

dat <- read_xlsx(here::here("data", "2021_chla-catalyst_data_all.xlsx"),
                 sheet = "qaqc") %>% 
    mutate(extracted = chla_ugl,
           sensor_rfu = chlorophyll_rfu,
           sensor_ugl = chl_fluor,
           month = lubridate::month(datetime_collected),
           season = case_when(month %in% c(12, 1, 2)  ~ "Winter",
                              month %in% c(3, 4, 5)   ~ "Spring",
                              month %in% c(6, 7, 8)   ~ "Summer",
                              month %in% c(9, 10, 11) ~ "Fall")) %>% 
    select(reserve = reserve_code,
           station = station_code,
           datetime_collected,
           month,
           season,
           method,
           extracted,
           sensor_rfu,
           temp,
           turb,
           fdom_qsu) %>% 
    mutate(station = case_when(reserve == "GND" ~ "gndblwq",
                               reserve == "SAP" ~ "sapnawq",
                               TRUE ~ station),
           reserve = case_when(station == "heew2wq" ~ "HEE_w2",
                               station == "WS" ~ "NIW_isco", # at NIW: OL is tank; WS is ISCO. These should not be pooled.
                               TRUE ~ reserve),
           station = case_when(station == "OL" ~ "niwolwq",
                               station == "WS" ~ "niwwswq",
                               TRUE ~ station)) %>% 
    drop_na() %>% 
    mutate(across(c(reserve, month, season),
                  as.factor),
           season = forcats::fct_relevel(season, c("Winter", "Spring", "Summer", "Fall")),
           rfu_temp_corrected = sensor_rfu / (1 + global_rho*(temp-20)))

# save(dat, file = here::here("data", "subset_no_NA.RData"))
# uncommented and ran the above, 2/16/22

# read back in later with:
# load(here::here("data", "subset_no_NA.RData"))
```

Quick gut-check on temp-corrected RFU:  

```{r}
ggplot(dat, aes(x = sensor_rfu, y = rfu_temp_corrected)) +
    geom_point(size = 2, alpha = 0.4, aes(color = reserve)) +
    geom_abline(slope = 1, intercept = 0, color = "gray40") +
    labs(x = "Sensor RFU",
         y = "Temperature-corrected RFU")
```

Points are generally close to the 1:1 line.  

## Summary  

### General  

```{r}
dat %>% group_by(reserve) %>% count(station) %>% knitr::kable()
summary(dat)
nrow(dat)
```


### Reserve by Month  

Seems like some of us didn't get the temporal spread we were wanting.  


```{r}
table(dat$reserve, dat$month)
```

```{r}
dat %>% 
    group_by(reserve, month, method) %>% 
    summarize(n = n()) %>% 
    ggplot(aes(x = method, y = n, col = reserve)) +
    geom_point(size = 2, alpha = 0.6, position = position_dodge2(width = 0.3)) +
    facet_wrap(~month, ncol = 4) +
    labs(title = "# Samples by Month, by Reserve") +
    theme_bw()
```

### Reserve by Season  

```{r}
table(dat$reserve, dat$season)
```

```{r}
dat %>% 
    group_by(reserve, season, method) %>% 
    summarize(n = n()) %>% 
    ggplot(aes(x = method, y = n, col = reserve)) +
    geom_point(size = 2, alpha = 0.6, position = position_dodge2(width = 0.3)) +
    facet_wrap(~season, ncol = 2) +
    labs(title = "# Samples by Season, by Reserve") +
    theme_bw()
```


Values by season and reserve.....

```{r}
dat %>% 
    group_by(reserve, season, method) %>% 
    ggplot(aes(x = method, y = extracted, col = reserve)) +
    geom_point(size = 2, alpha = 0.6, position = position_dodge2(width = 0.3)) +
    facet_wrap(~season, ncol = 2) +
    labs(title = "Extracted chl a by Season, by Reserve") +
    theme_bw()
```


```{r}
dat %>% 
    group_by(reserve, season, method) %>% 
    ggplot(aes(x = method, y = fdom_qsu, col = reserve)) +
    geom_point(size = 2, alpha = 0.6, position = position_dodge2(width = 0.3)) +
    facet_wrap(~season, ncol = 2) +
    labs(title = "FDOM (qsu) by Season, by Reserve") +
    theme_bw()
```


### Correlations between predictors?  

```{r, fig.width = 8, fig.height = 8}
dat %>% 
    select(reserve, month, season, method, extracted, sensor_rfu, temp, turb, fdom_qsu) %>% 
    plot()
```

