---
output: 
    html_document
---


```{r}
# library(readxl)
# library(dplyr)
# library(tidyr)
# library(ggplot2)
# library(patchwork)
# 
load(here::here("data", "subset_no_NA.RData"))

# original plot settings, for later
op <- par()
```

## Full Model with all interactions  

response ~ sensor_rfu + turb + fdom_qsu + sensor_rfu\*turb\*fdom_qsu + sensor_rfu\*turb + sensor_rfu\*fdom_qsu + turb\*fdom_qsu + temp + month + reserve    


### No transformation of response  

#### Fit Model  

```{r}
full <- lm(extracted ~ sensor_rfu + turb + fdom_qsu + sensor_rfu*turb*fdom_qsu + sensor_rfu*turb + sensor_rfu*fdom_qsu + turb*fdom_qsu + temp + month + reserve, data = dat)

# diagnostics
df_untransf <- broom::glance(full)
```

#### Plot residuals  

```{r}
par(mfrow = c(2, 2),
    mar = c(2, 2, 2, 1))
plot(full)
```


### Square-root transformation of response  

#### Fit Model  

```{r}
full_sqrt <- lm(sqrt(extracted) ~ sensor_rfu + turb + fdom_qsu + sensor_rfu*turb*fdom_qsu + sensor_rfu*turb + sensor_rfu*fdom_qsu + turb*fdom_qsu + temp + month + reserve, data = dat)

# diagnostics  
df_sqrt <- broom::glance(full_sqrt)
```


#### Plot residuals  

```{r}
par(mfrow = c(2, 2),
    mar = c(2, 2, 2, 1))
plot(full_sqrt)
```


### Fourth-root transformation of response  

Similar to a natural-log transformation, but log(0) doesn't exist so I'd like to avoid that transformation.    

#### Fit Model  

```{r}
full_frthrt <- lm(extracted^0.25 ~ sensor_rfu + turb + fdom_qsu + sensor_rfu*turb*fdom_qsu + sensor_rfu*turb + sensor_rfu*fdom_qsu + turb*fdom_qsu + temp + month + reserve, data = dat)

# diagnostics
df_frthrt <- broom::glance(full_frthrt)
```

#### Plot residuals  

```{r}
par(mfrow = c(2, 2),
    mar = c(2, 2, 2, 1))
plot(full_frthrt)
```


### Log transformation of response  

Because this is common and I got asked about it in a Data Team meeting.    

#### Fit Model  

```{r}
full_log10 <- lm(log10(extracted + 0.1) ~ sensor_rfu + turb + fdom_qsu + sensor_rfu*turb*fdom_qsu + sensor_rfu*turb + sensor_rfu*fdom_qsu + turb*fdom_qsu + temp + month + reserve, data = dat)

# diagnostics
df_log10 <- broom::glance(full_log10)
```

#### Plot residuals  

```{r}
par(mfrow = c(2, 2),
    mar = c(2, 2, 2, 1))
plot(full_log10)
```



## Diagnostics and predictions for all  

Predictions use exponents to get them back on the original scale.  

```{r}
dat_preds <- dat %>% 
    mutate(preds_full_untr = fitted.values(full),
           preds_full_sqrt = (fitted.values(full_sqrt))^2,
           preds_full_frthrt = (fitted.values(full_frthrt))^4,
           preds_full_log10 = 10^(fitted.values(full_log10)),
           
           abs_dev_untr = abs(preds_full_untr - extracted),
           abs_dev_sqrt = abs(preds_full_sqrt - extracted),
           abs_dev_frthrt = abs(preds_full_frthrt - extracted),
           abs_dev_log10 = abs(preds_full_log10 - extracted))

fits <- tribble(
    ~"transformation", ~"R2", ~"AdjR2", ~"MedAbsDev",
    "none", df_untransf$r.squared, df_untransf$adj.r.squared, median(dat_preds$abs_dev_untr),
    "square-root", df_sqrt$r.squared, df_sqrt$adj.r.squared, median(dat_preds$abs_dev_sqrt),
    "fourth-root", df_frthrt$r.squared, df_frthrt$adj.r.squared, median(dat_preds$abs_dev_frthrt),
    "log10(x+0.1)", df_log10$r.squared, df_log10$adj.r.squared,
    median(dat_preds$abs_dev_log10)
)
```

### Diagnostics summary  

```{r}
knitr::kable(fits, caption = "Diagnostics for model fit", digits = 3)
```

### Observed vs. Predicted Plots  

Units are all ug/L.  

The black line on the graph is the 1:1 line: we want points to be close to this line.  

```{r}
p1 <- ggplot(dat_preds, aes(x = preds_full_untr, y = extracted, color = reserve)) +
    geom_point(size = 2, alpha = 0.4) +
    geom_abline(slope = 1, intercept = 0) +
    theme_bw() +
    labs(title = "Untransformed Model: Observed vs. Predicted Chlorophyll a",
         x = "Predicted",
         y = "Extracted")

p2 <- ggplot(dat_preds, aes(x = preds_full_sqrt, y = extracted, color = reserve)) +
    geom_point(size = 2, alpha = 0.4) +
    geom_abline(slope = 1, intercept = 0) +
    theme_bw() +
    labs(title = "Square-root transformation: Observed vs. Predicted Chlorophyll a",
         x = "Predicted",
         y = "Extracted")

p3 <- ggplot(dat_preds, aes(x = preds_full_frthrt, y = extracted, color = reserve)) +
    geom_point(size = 2, alpha = 0.4) +
    geom_abline(slope = 1, intercept = 0) +
    theme_bw() +
    labs(title = "Fourth-root transformation: Observed vs. Predicted Chlorophyll a",
         x = "Predicted",
         y = "Extracted")

p4 <- ggplot(dat_preds, aes(x = preds_full_log10, y = extracted, color = reserve)) +
    geom_point(size = 2, alpha = 0.4) +
    geom_abline(slope = 1, intercept = 0) +
    theme_bw() +
    labs(title = "Log10 transformation: Observed vs. Predicted Chlorophyll a",
         x = "Predicted",
         y = "Extracted")
```

```{r, fig.width = 6, fig.height = 7}
(p1 + p2) / (p3 + p4) + plot_layout(guides = "collect")
```


Square-root transformation seems to do the best walking-the-line to get:  

+  no negative predictions for extracted chlorophyll  
+  generally okay linear fit without extreme prediction outliers or harsh transformation  

