---
title: "08_playing_with_graphics"
author: "Kim Cressman"
date: "12/10/2021"
output: 
    html_document:
        code_folding: hide
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

# Setup  

```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(lmerTest)
library(MuMIn)

load(here::here("data", "subset_no_NA.RData"))

# original plot settings, for later
op <- par()

# options, for MuMIn operations
# will prevent fitting sub-models to different datasets (from help file)
options(na.action = "na.fail")
```

One manipulation of data frame so I can code this up and make a final decision on a transformation later: generate a column called `response` that will be used in the models. If we decide on a square-root transformation, response will be calculated as `sqrt(extracted)`. That's what I'll work with for now but it can be changed to `extracted^(1/3)` or something different, and we should still be able to generate all relevant outputs.  

```{r}
dat$response <- sqrt(dat$extracted)
```

# Make a lot of predictions and graph them  

For each reserve  

## Global models and predictions  

```{r}
glob_randInt <- lmer(response ~ sensor_rfu*(temp + turb*fdom_qsu + season) + (1|reserve),
                    data = dat)

glob_randSlope <- lmer(response ~ sensor_rfu*(temp + turb*fdom_qsu + season) + (sensor_rfu|reserve),
             data = dat)

glob_ols_noReserve <- lm(response ~ sensor_rfu*(temp + turb*fdom_qsu + season),
                         data = dat)


dat2 <- dat %>% 
    mutate(predicted_ols = (predict(glob_ols_noReserve, newdata = .))^2,
           predicted_randInt_noRes = (predict(glob_randInt, newdata = ., re.form = ~0))^2,
           predicted_randInt_knownRes = (predict(glob_randInt, newdata = .))^2,
           predicted_randSlope_noRes = (predict(glob_randInt, newdata = ., re.form = ~0))^2,
           predicted_randSlope_knownRes = (predict(glob_randSlope, newdata = .))^2)
```

## Reserve level, and graphs  

```{r}

reserves <- unique(dat$reserve)

for(i in seq_along(reserves)){
    res <- dat2 %>% 
        filter(reserve == reserves[i]) %>% 
        droplevels()
    # DROPPING LEVELS CAN BE DANGEROUS
    
    res_ols <- lm(response ~ sensor_rfu*(temp + turb*fdom_qsu + season),
                  data = res)

    res$preds_ols <- (predict(res_ols, newdata = res))^2
    
    print(
        ggplot(res, aes(y = extracted)) +
            geom_abline(slope = 1, intercept = 0) +
            geom_point(aes(x = predicted_randInt_noRes,
                           color = "Intercept", shape = "unknown"),
                       size = 2, alpha = 0.4) +
            geom_point(aes(x = predicted_randInt_knownRes,
                           color = "Intercept", shape = "known"),
                       size = 2, alpha = 0.4) +
            geom_point(aes(x = predicted_randSlope_noRes,
                           color = "Slope", shape = "unknown"),
                       size = 2, alpha = 0.4) +
            geom_point(aes(x = predicted_randSlope_knownRes,
                           color = "Slope", shape = "known"),
                       size = 2, alpha = 0.4) +
            geom_point(aes(x = preds_ols,
                           color = "OLS", shape = "known"),
                       size = 2, alpha = 0.4) +
            labs(title = reserves[i],
                 x = "Predicted",
                 y = "Observed",
                 color = "Random",
                 shape = "Reserve as predictor")
    )
}

```

